---
title: "Vignettes of the `influential`"
output: 
  rmarkdown::github_document:
    toc: true
    toc_depth: 2
    keep_html: true
vignette: >
  %\VignetteIndexEntry{influential vignettes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(tibble.print_min = 4L, tibble.print_max = 4L)
set.seed(60)
```

## Overview

The influential package contains several functions that could be categorized into four groups
according to their purpose:

* Network reconstruction
* Calculation of centrality measures
* Assessment of the association of centrality measures
* Identification of the most `influential` network nodes (hubs)
    
********

```{r setup}
library(influential)
```

## Network reconstruction

Two functions have been obtained from the `igraph` package for the reconstruction of networks.

- ### From a data frame
    In the data frame the first and second columns should be composed of source and target nodes.
    
    A sample appropriate data frame is brought below:
    
```{r, echo=FALSE}
knitr::kable(head(coexpression.data))
```
    
This is a co-expression dataset obtained from [PMID: 31211495](https://www.ncbi.nlm.nih.gov/pubmed/31211495).

```{r g_dataframe}
MyData <- coexpression.data        # Preparing the data

My_graph <- graph_from_data_frame(d=MyData)        # Reconstructing the graph
```

If you look at the class of `My_graph` you should see that it has an `igraph` class:
```{r}
class(My_graph)
```

********

- ### From an adjacency matrix

A sample appropriate data frame is brought below:

```{r, echo=FALSE}
knitr::kable(head(coexpression.adjacency)[,1:10])
```

```{r g_adj}
MyData <- coexpression.adjacency        # Preparing the data

My_graph <- graph_from_adjacency_matrix(MyData)        # Reconstructing the graph
```

********

- ### Network vertices

Network vertices (nodes) are required in order to calculate their centrality measures.
Thus, before calculation of network centrality measures we need to obtain the name of required network
vertices. To this end, we use the `V` function, which is obtained from the `igraph` package.

```{r Vertices}
MyData <- coexpression.data        # Preparing the data

My_graph <- graph_from_data_frame(MyData)        # Reconstructing the graph

My_graph_vertices <- V(My_graph)        # Extracting the vertices
```

********

## Calculation of centrality measures

- ### Degree centrality
Degree centrality is the most commonly used local centrality measure which could be calculated
via the `degree` function obtained from the `igraph` package.

```{r DC}
MyData <- coexpression.data        # Preparing the data

My_graph <- graph_from_data_frame(MyData)        # Reconstructing the graph

GraphVertices <- V(My_graph)        # Extracting the vertices

My_graph_degree <- degree(My_graph, v = GraphVertices, normalized = FALSE) # Calculating degree centrality
```

Degree centrality could be also calculated for *directed* graphs via specifying the `mode` parameter.

- ### Betweenness centrality
Betweeneess centrality, like degree centrality, is one of the most commonly used centrality measures but is representative of the global centrality of a node. This centrality metric could also be  calculated 
using a function obtained from the `igraph` package.

```{r BC}
MyData <- coexpression.data        # Preparing the data

My_graph <- graph_from_data_frame(MyData)        # Reconstructing the graph

GraphVertices <- V(My_graph)        # Extracting the vertices

My_graph_betweenness <- betweenness(My_graph, v = GraphVertices,    # Calculating betweenness centrality
                                    directed = FALSE, normalized = FALSE)
```

Degree centrality could be also calculated for *directed* and/or *weighted* graphs via specifying the `directed` and `weights` parameters, respectively.

- ### Neighborhood connectivity

Neighborhood connectivity is one of the other important centrality measures that reflect the (semi-) local
centrality of a node. This centrality measure was first represented in a [Science paper](https://www.ncbi.nlm.nih.gov/pubmed/11988575) in 2002 and is for the first time calculable in R
environment via the `influential` package.

```{r NC}
MyData <- coexpression.data        # Preparing the data

My_graph <- graph_from_data_frame(MyData)        # Reconstructing the graph

GraphVertices <- V(My_graph)        # Extracting the vertices

neighrhood.co <- neighborhood.connectivity(graph = My_graph,    # Calculating neighborhood connectivity
                                           vertices = GraphVertices,
                                           mode = "all")
```

Neighborhood connectivity could be also calculated for *directed* graphs via specifying the `mode` parameter.

********

## Assessment of the association of centrality measures

- ### Conditional probability of deviation from means

The function `cond.prob.analysis` assesses the conditional probability of deviation of two
centrality measures (or any other two continuous variables) from their corresponding means in oposite directions.

```{r cond.prob}
MyData <- centrality.measures        # Preparing the data

My.conditional.prob <- cond.prob.analysis(data = MyData,       # Assessing the conditional probability
                                          nodes.colname = "name",
                                          Desired.colname = "BetweennessCentrality",
                                          Condition.colname = "NeighborhoodConnectivity")

print(My.conditional.prob)
```

- As you can see in the results, the whole data is also randomly splitted into half in order 
to further test the validity of conditional probability assessments.
- *The higher the conditional probability the more two centrality measures behave in contrary manners*.

********

- ### Nature of association (considering dependent and independent)

The function `double.cent.assess` could be used to automaticelly assess both the distribution
mode of centrality measures (two continuous variables) and the nature of their association. The analyses done through this formula are as follows:

1. **Normality assessment**:
    - Variables with **lower than** 5000 observations: *Shapiro-Wilk test*
    - Variables with **over** 5000 observations: *Anderson-Darlin test*
    <br><br>
2. **Assessment of non-linear/non-monotonic correlation**:
    - *Non-linearity assessment*: Fitting a generalized additive model (GAM) with integrated smoothness approximations 
    using the `mgcv` package
        <br><br>
    - *Non-monotonicity assessment*: Comaring the squared coefficients of the correlation based on Spearman's rank 
    correlation analysis and ranked regression test with non-linear splines.
        - Squared coefficient of Spearman's rank correlation **>** R-squared ranked regression with non-linear splines: *Monotonic*
        - Squared coefficient of Spearman's rank correlation **<** R-squared ranked regression with non-linear splines: *Non-monotonic*
    <br><br>
3. **Dependemnce assessment**:
    - *Hoeffding’s independence test*: Hoeffding's test of independence is a test based on the population 
    measure of deviation from independence which computes a D Statistics ranging from -0.5 to 1: Greater D values 
    indicate a higher dependence between variables.
    - *Descriptive non-linear non-parametric dependence test*: This assessment is based on non-linear non-parametric 
    statistics (NNS) which outputs a dependence value ranging from 0 to 1. For further details please refer to [NNS: Nonlinear Nonparametric Statistics](https://cran.r-project.org/package=NNS): Greater values indicate a higher dependence between variables.
        <br><br>
4. **Correlation assessment**:
As the correlation between most of the centrality measures follows a non-monotonic form, this part of the assessment
is done based on the non-linear non-parametric statistics (NNS) which itself calculates the correlation based on partial moments and outputs a correlation value ranging from -1 to 1. For further details please refer to [NNS: Nonlinear Nonparametric Statistics](https://cran.r-project.org/package=NNS).
        <br><br>
5. **Assessment of conditional probability of deviation from means**
This step assesses the conditional probability of deviation of two centrality measures (or any other two continuous variables) from their corresponding means in oposite directions.
    - The independent centrality measure (variable) is considered as the condition variable and the other as the desired one.
    - As you will see in the results, the whole data is also randomly splitted into half in order 
to further test the validity of conditional probability assessments.
    - *The higher the conditional probability the more two centrality measures behave in contrary manners*.

```{r double.cent.assess}
MyData <- centrality.measures        # Preparing the data

My.metrics.assessment <- double.cent.assess(data = MyData,       # Association assessment
                                            nodes.colname = "name",
                                            dependent.colname = "BetweennessCentrality",
                                            independent.colname = "NeighborhoodConnectivity")

print(My.metrics.assessment)
```

**Note**: It should also be noted that as a single regression line does not fit all models with a 
certain degree of freedom, based on the size and correlation mode of the variables provided, this function might return an error due to incapability of running step 2. In this case, you may follow each step manually or as an alternative run the other function named `double.cent.assess.noRegression` which does not perform any regression test and consequently it is not required to determine the dependent and independent variables.

********

- ### Nature of association (without considering dependence direction)

The function `double.cent.assess.noRegression` could be used to automaticelly assess both the distribution
mode of centrality measures (two continuous variables) and the nature of their association. The analyses done through this formula are as follows:

1. **Normality assessment**:
    - Variables with **lower than** 5000 observations: *Shapiro-Wilk test*
    - Variables with **over** 5000 observations: *Anderson-Darlin test*
    <br><br>
2. **Dependemnce assessment**:
    - *Hoeffding’s independence test*: Hoeffding's test of independence is a test based on the population 
    measure of deviation from independence which computes a D Statistics ranging from -0.5 to 1: Greater D values 
    indicate a higher dependence between variables.
    - *Descriptive non-linear non-parametric dependence test*: This assessment is based on non-linear non-parametric 
    statistics (NNS) which outputs a dependence value ranging from 0 to 1. For further details please refer to [NNS: Nonlinear Nonparametric Statistics](https://cran.r-project.org/package=NNS): Greater values indicate a higher dependence between variables.
        <br><br>
3. **Correlation assessment**:
As the correlation between most of the centrality measures follows a non-monotonic form, this part of the assessment
is done based on the non-linear non-parametric statistics (NNS) which itself calculates the correlation based on partial moments and outputs a correlation value ranging from -1 to 1. For further details please refer to [NNS: Nonlinear Nonparametric Statistics](https://cran.r-project.org/package=NNS).
        <br><br>
4. **Assessment of conditional probability of deviation from means**
This step assesses the conditional probability of deviation of two centrality measures (or any other two continuous variables) from their corresponding means in oposite directions.
    - The `centrality2` variable is considered as the condition variable and the other (`centrality1`) as the desired one.
    - As you will see in the results, the whole data is also randomly splitted into half in order 
to further test the validity of conditional probability assessments.
    - *The higher the conditional probability the more two centrality measures behave in contrary manners*.

```{r double.cent.assess.noRegr., eval=FALSE}
MyData <- centrality.measures        # Preparing the data

My.metrics.assessment <- double.cent.assess.noRegression(data = MyData,       # Association assessment
                                                         nodes.colname = "name",
                                                         centrality1.colname = "BetweennessCentrality",
                                                         centrality2.colname = "NeighborhoodConnectivity")

print(My.metrics.assessment)
#> $Summary_statistics
#>         BetweennessCentrality NeighborhoodConnectivity
#> Min.              0.000000000                   1.2000
#> 1st Qu.           0.000000000                  66.0000
#> Median            0.000000000                 156.0000
#> Mean              0.005813357                 132.3443
#> 3rd Qu.           0.000340000                 179.3214
#> Max.              0.529464720                 192.0000
#> 
#> $Normality_results
#>                               p.value
#> BetweennessCentrality    1.415450e-50
#> NeighborhoodConnectivity 9.411737e-30
#> 
#> $Centrality1_Normality
#> [1] "Non-normally distributed"
#> 
#> $Centrality2_Normality
#> [1] "Non-normally distributed"
#> 
#> $HoeffdingD_Statistic
#>         D_statistic P_value
#> Results  0.01770279   1e-08
#> 
#> $Dependence_Significance
#>                       Hoeffding
#> Results Significantly dependent
#> 
#> $NNS_dep_results
#>         Correlation Dependence
#> Results  -0.7948106  0.8647164
#> 
#> $ConditionalProbability
#> [1] 55.35386
#> 
#> $ConditionalProbability_split.half.sample
#> [1] 55.68163
```

********

## Identification of the most `influential` network nodes (hubs)

**IHS**
    : `IHS` is the first integrative method for the identification of network hubs. The `IHS` formula integrates three
    network centrality meassure including degree centrality, betweenness centrality, and neighborhood connectivity in
    such a way that both synergize their effects and remove their biases.
    
```{r ihs}
MyData <- centrality.measures        # Preparing the data

My.vertices.IHS <- ihs(DC = centrality.measures$Degree,       # Calculation of IHS
                       BC = centrality.measures$BetweennessCentrality,
                       NC = centrality.measures$NeighborhoodConnectivity)

print(head(My.vertices.IHS))
```

