---
title: "Vignettes of the `influential`"
output: 
  rmarkdown::github_document:
    toc: true
    toc_depth: 2
    keep_html: true
vignette: >
  %\VignetteIndexEntry{influential vignettes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(tibble.print_min = 4L, tibble.print_max = 4L)
set.seed(60)
```

## Overview

The influential package contains several functions that could be categorized into four groups
according to their purpose:

* Network reconstruction
* Calculation of centrality measures
* Assessment of the association of centrality measures
* Identification of the most `influential` network nodes (hubs)
    
********

```{r setup}
library(influential)
```

## Network reconstruction

Two functions have been obtained from the `igraph` package for the reconstruction of networks.

- ### From a data frame
    In the data frame the first and second columns should be composed of source and target nodes.
    
    A sample appropriate data frame is brought below:
    
```{r, echo=FALSE}
knitr::kable(head(coexpression.data))
```
    
This is a co-expression dataset obtained from [PMID: 31211495](https://www.ncbi.nlm.nih.gov/pubmed/31211495).

```{r}
MyData <- coexpression.data
My_graph <- graph_from_data_frame(d=MyData)
```

If you look at the class of `My_graph` you should see that it has an `igraph` class:
```{r}
class(My_graph)
```

********

- ### From an adjacency matrix

A sample appropriate data frame is brought below:

```{r, echo=FALSE}
knitr::kable(head(coexpression.adjacency)[,1:10])
```

```{r}
MyData <- coexpression.adjacency
My_graph <- graph_from_adjacency_matrix(MyData)
```

********

## Calculation of centrality measures

- ### Degree centrality

- ### Betweenness centrality

- ### Neighborhood connectivity



********

## Assessment of the association of centrality measures


## Identification of the most `influential` network nodes (hubs)
